---
title: "blocks"
output: github_document
date: "2023-09-04"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(caret)
library(broom)
library(ggrepel)
```

i'm lazy and i don't have block counts lying around so we're going to fudge block counts with (mins played) x (block %). close enough for the demonstration here.

```{r}
df = read_csv('~/Dropbox/funfund/data/vorp.csv') %>%
  select(
    season=Season,
    player=Player,
    pos=Pos,
    age=Age,
    team=Tm,
    block_pct=`BLK%`,
    mins=MP
  ) %>%
  mutate(
    blocks_fudged=round(block_pct * mins / 100),
    z_score = (blocks_fudged - mean(blocks_fudged)) / sd(blocks_fudged)
  ) %>%
  filter(season==2023)
```

```{r}
hist(df$blocks_fudged, breaks=30)
```

```{r}
hist(df$z_score)
```

```{r}
lambda_hat = mean(df$blocks_fudged)
N = nrow(df)
```

if poisson distributed

```{r}
hist(rpois(N, lambda_hat))
```

so that doesn't look great

let's try exponential dist

```{r}
hist(rexp(N, 1/lambda_hat))
```
well that looks great!!

```{r}
hist(df$blocks_fudged, col=alpha('purple', 0.5), main='purple = data, yellow = model', xlab='blocks')
hist(rexp(N, 1/lambda_hat), col=alpha('yellow', 0.5), add=T)
```


let's compute empirical "exponential scores"

these are going to be the percentile of the observation from 0 to 1

```{r}
percentiles = pexp(df$blocks_fudged, rate=1/lambda_hat)
percentiles = ifelse(percentiles == 0, 0.0002, percentiles)
```

```{r}
hist(percentiles)
```

now clip these values and pretend these come from a normal dist

we still have a zeros problem but rad! this looks great

```{r}
hist(qnorm(percentiles))
```

let's go look at the top players

```{r}
df$percentiles = percentiles
df$pseudo_z_score = qnorm(percentiles)
```

```{r}
df %>%
  arrange(-percentiles) %>%
  select(player, pos, block_pct, pseudo_z_score, z_score)
```

```{r}
plot(pseudo_z_score ~ z_score, data=df)
```

